hbm_paper_results

### 2024-01-16 paper_prop_results.Rmd
troubleshooting

*need to add w/in feature sex-bias testing, having issues with cubic*

@@

```{r, message=FALSE}
ratio.merge.df <- full_join(ratio_raw_pred.df, ratio.df)
unique(ratio.merge.df$dataset)
dim(ratio_raw_pred.df)
dim(ratio.merge.df)

ratio.merge.df <- ratio.merge.df %>%
  dplyr::select(!File_ID) %>%
  mutate(dataset=gsub("_predictions.csv","", dataset))
```

@@

```{r}
# df_list <- c()
# for (p in unique(ratio.merge.df$prop)) {
#   df <- ratio.merge.df %>%
#     dplyr::filter(prop == p)
#   df_list[p] <- list(df)
# }
# length(df_list) == length(unique(ratio.merge.df$prop))
```

define function to remove extremes
```{r}
rm_ext_cents <- function(og.data, feat_list){
  df_ext_list <- c()
  count <- 1
  for(pheno in feat_list) {
    list <- c(pheno, paste0("diff_", pheno), paste0("abs.diff_", pheno))

    df <- og.data %>%
      dplyr::select(participant, sex, age_days, sim.site, Source_File, dataset, prop, all_of(list)) %>% #get correct cols
      group_by(participant) %>% 
      dplyr::filter(!any((get(pheno) < 0.05 | get(pheno) > 0.95) & dataset == "raw")) %>% #drop ppts if raw centile is <0.05 or >0.95
      ungroup() %>%
      dplyr::filter(dataset != "raw") #now drop raw vals
    df_ext_list[count] <- list(df)
    count <- count + 1
  }
  stopifnot(length(df_ext_list) == length(pheno_list))
  df.no_ext <- Reduce(full_join, df_ext_list)
  return(df.no_ext)
}
```

apply fun

```{r}
ratio_no_ext_list <- mclapply(df_list, rm_ext_cents, feat_list=pheno_list, mc.preschedule = FALSE)
```

>>>>

*need to add w/in feature sex-bias testing*

@@

```{r}
ratio.no_ext_cent <- fread(file=paste0(data_path, "/ukb_ratio/no_ext_cent.csv"))
```

@@

```{r}
ratio.no_ext_cent_list <- c()
for (p in unique(ratio.no_ext_cent$prop)){
  df <- ratio.no_ext_cent %>%
    dplyr::filter(prop == p)
  ratio.no_ext_cent_list <- c(ratio.no_ext_cent_list, list(df))
}
rm(ratio.no_ext_cent)
length(ratio.no_ext_cent_list)
dim(ratio.no_ext_cent_list[[1]]) == dim(ratio.no_ext_cent_list[[9]]) #expect TRUE
sum(unique(ratio.no_ext_cent_list[[1]]$participant) == unique(ratio.no_ext_cent_list[[9]]$participant)) #expect low # matches
```

Test between-combat abs. centile diffs - probably faster if i just run this back on cubic

```{r}
noext.prop_abs.cent_t.tests <- mclapply(ratio.no_ext_cent_list, centile.t.tests, feature_list=pheno_abs.diff.list, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
names(noext.prop_abs.cent_t.tests) <- n_prop_list
noext.prop_abs.cent_t.tests_df <- bind_rows(noext.prop_abs.cent_t.tests, .id = "column_label")
noext.prop_abs.cent_t.tests_df
```

*need to add w/in feature sex-bias testing*

## Z-SCORES

Do z score errors change based on combat version?

```{r}
prop_abs.z_t.tests <- mclapply(ratio_list, centile.t.tests, feature_list=z.pheno_abs.diff.list, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
names(prop_abs.z_t.tests) <- n_prop_list
prop_abs.z_t.tests_df <- bind_rows(prop_abs.z_t.tests, .id = "column_label")
```

Visualize 
```{r}
ggplot(prop_abs.z_t.tests_df) +
  geom_col(aes(x=column_label, y=n_sig_regions, fill=column_label)) +
  geom_hline(yintercept = 208, linetype='longdash') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none") +
  facet_wrap(~comp)
```

prop.global_z.abs.diffs <- ratio.df %>%
  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("abs.diff_", vol_list_global, ".z")) %>%
  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff")

prop.global_z_errs_plot <- prop.global_z.abs.diffs %>%
  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
         abs.diff = abs.diff*100) %>% #for visuals
  ggplot() +
  geom_violin(aes(x=prop, y=abs.diff, fill=dataset, color=dataset), scale="count") +
  theme_linedraw(base_size = 11) +
  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y="Magnitude Z Error", x = "% Females in Imbalanced Study Site") +
  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_discrete(labels=percent_f) +
  facet_wrap(~ pheno) +
  ggtitle("Magnitude errors in z scores of global features across ComBat pipelines")
prop.global_z_errs_plot
#ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
```

*need to add w/in feature sex-bias testing*

```{r}
ratio.no_ext_cent <- fread(file=paste0(data_path, "/ukb_ratio/no_ext_cent.csv"))
```

```{r}
ratio.no_ext_cent_list <- c()
for (p in unique(ratio.no_ext_cent$prop)){
  df <- ratio.no_ext_cent %>%
    dplyr::filter(prop == p)
  ratio.no_ext_cent_list <- c(ratio.no_ext_cent_list, list(df))
}
rm(ratio.no_ext_cent)
length(ratio.no_ext_cent_list)
dim(ratio.no_ext_cent_list[[1]]) == dim(ratio.no_ext_cent_list[[9]]) #expect TRUE
sum(unique(ratio.no_ext_cent_list[[1]]$participant) == unique(ratio.no_ext_cent_list[[9]]$participant)) #expect low # matches
```

Test between-combat abs. centile diffs - probably faster if i just run this back on cubic

```{r}
noext.prop_abs.cent_t.tests <- mclapply(ratio.no_ext_cent_list, centile.t.tests, feature_list=pheno_abs.diff.list, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
names(noext.prop_abs.cent_t.tests) <- n_prop_list
noext.prop_abs.cent_t.tests_df <- bind_rows(noext.prop_abs.cent_t.tests, .id = "column_label")
noext.prop_abs.cent_t.tests_df
```

*need to add w/in feature sex-bias testing*
